---
title: "Creating_functions_in_R"
author: "Christina Minions"
date: "2/13/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#work with some data in degrees F 

- convert to C  

Sourcing functions within an rscript is handy - do not have to copy and paste functions into a new script. Once you have created the function within an R script, you can just call the script using the source function, and the function(s) that you created will be loaded into your R environment 

```{r}
source("Convert.R") #after putting the functions in an Rscript and saving, you can call the functions at the start of code 
airtemps <-c(212, 100, 32, 64, 0, -20)


airtemp1 <- (airtemps[1]-32)* 5/9
```




#Write a function to do that

```{r}
fahr_to_celsius <- function(fahr) {
  celsius <- (fahr-32) * 5/9   #assigning this value to a object called celsius 
  return(celsius)   #use the return function to be explicit about what value is going to be the ouput
}

#once you run this, the function is in the R environment and avaible for use! 


fahr_to_celsius(airtemps[1]) #runs function for just first column 

celsius_temps <- fahr_to_celsius(airtemps) #runs function for all objects in this vector 
celsius_temps

```

#write a function to go C to F 
Creating opposite function to return back F 
```{r}

Cels_to_Fahr <- function(Cels) {
  FahrenH <- (celsius_temps) * 9/5 + 32   #assigning this value to a object called celsius 
  return(FahrenH)   #use the return function to be explicit about what value is going to be the ouput
}


Cels_to_Fahr(celsius_temps) #runs function for just first column 

Back_to_FAHR <- Cels_to_Fahr(celsius_temps) #runs function for all objects in this vector 

Back_to_FAHR


```



# Show how to document Functions 

```{r}

#'Coverts numeric temperature data from degrees Fahrenhiet to Degrees Celsius
#'
#'
#'@param fahr the value in Fahrenhiet to be converted
#'@return the converted value in Celsius 
fahr_to_celsius <- function(fahr) {
  celsius <- (fahr-32) * 5/9   
  return(celsius)   
}



```







