---
title: "Data_Remformatting"
author: "Christina Minions"
date: "2/13/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,warning = FALSE, message = FALSE} 
#Load in the Libraries 
library(dplyr)
library(tidyr)
```

# Reformat the Catch Data 

*Remove the "all" column 
*Create a species column 
    - move from wide to long 
*General QA 

###This is the citation for the dataset (https://knb.ecoinformatics.org/view/df35b.304.2)
Mike Byerly. Alaska commercial salmon catches by management region (1886- 1997). urn:node:GOA. df35b.304.2.

```{r}
catch_original <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1", method = "libcurl"),
                    stringsAsFactors = FALSE)
head(catch_original)
```

### Removing the 'ALL' column and the 'notesRegCode' column from the dataframe
  using a function called 'select'
  command + shift + m is the shortcut for the pipe function %>% 
  To make data into the "Long" format, using the function 'gather'
  
This is gathering the data over every column except for the data within the columns of Region and Year  
    * gather(key = "species", value = "catch", -Region, -Year)   
  
  
```{r}

catch_long <- catch_original %>% 
    #select(Region, Year, Chinook, Sockeye, Coho, Pink, Chum) #This selects the columns you want
    select(-All,-notesRegCode) %>% #This de-selects the columns you don't want
    gather(key = "species", value = "catch", -Region, -Year) 

head(catch_long)
```

###For some reason if we wanted to go back and have the data in a wide format, use the function 'spread' 
```{r}
catch_wide <- catch_long %>%  
    spread(key = "species", value = "catch")
```

##Clean up data 

*rename column 'Catch' to 'catch_thousands' 
*change catch column from character to numeric 
*Create new catch column in units: `#` of fish 

For the rename function the new name of the column is equal to the name of the old column  
The mutate function will allow you to manipulate the values of a column 

 mutate(catch_thousands = ifelse(catch_thousands %in% c("I","M","Z"), 1, catch_thousands))

  ^can use this code to select and replace multiple different values with the same value. So if for some reason the dataframe had I, M and Z within in it and you wanted to replace all with 1. 
```{r}

catch_clean <- catch_long %>% 
  rename(catch_thousands = catch) %>% 
  mutate(catch_thousands = ifelse(catch_thousands == "I", 1, catch_thousands)) %>%  
  mutate(catch_thousands = as.numeric(catch_thousands)) %>% 
  mutate(catch = catch_thousands * 1000) %>% 
  select(-catch_thousands)

  
head(catch_clean)  
```

##Look for where as.numeric failed 
  Run this code to investigate.  
  For some reason the value got transformed from a 1 to an I when the files were converted after   looking at the original data file.   
  Need to change the I back to a 1 within R. DO NOT CHANGE the original data file 
  
  "mutate(catch_thousands = ifelse(catch_thousands == "I", 1, catch_thousands))" 
    * For this line of code, what it means is that if in the datafram "catch_thousands", any         value is equal to I, the value will be replaced with 1. If values do not equal I, the          original value of catch_thousands for the row will be used. 
    
```{r, eval = F, echo = F}
#Wanted to keep this code for my reference, but with these additions, it will not be in the output and R will not run it. 
test <- as.numeric(catch_long$catch) 
i <- which(is.na(test) == TRUE) # which values in test are equal to NA - shows 401 (this row)
catch_long[i,]

```

##Split - Apply - Combine 

    *calculate mean catch by species 
    
```{r}

species_mean <- catch_clean %>% 
  group_by(species, Region) %>%  #Grouping by species and region 
  summarize(catch_mean = mean(catch), #creating a column of the mean, using the values from catch.
            num_obs = n())            #Number of observations used to calculate the mean 

head(species_mean)


```
 by grouping by species and region you are calculating the mean for each species for each region. 

```{r}
species_max <- catch_clean %>% 
  group_by(species, Region) %>%  #Grouping by species and region 
  summarize(catch_max = max(catch), #creating a column of the mean, using the values from catch.
            num_obs = n())            #Number of observations used to calculate the mean 

head(species_max)
```


```{r}
year_mean <- catch_clean %>%  
  group_by(Year) %>%  
  summarize(catch_mean = mean(catch)) %>% 
  arrange(Year) %>%  # this is arrange the data in ascending order by year, for numeric values put a -for decesending 
  filter(Year >= 1990) #filters the data based on this condition. If we were only interested in years later than 1990 inclusive.  #to get the middle of a section: filter(Year <= 1990 & Year >= 1960). 
#use | for "or". so this line of code will give you year less than or equal to 1990 and greater or equal to 1960 

head(year_mean)
```

### Load the Region Defs File 

```{r}
region_defs <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.303.1",
                            method = "libcurl"),
                        stringsAsFactors = FALSE) %>% 
    select(code, mgmtArea)

head(region_defs)


```



###Joing two dataframes 
```{r}
catch_joined <- left_join(catch_clean, region_defs, by = c("Region" = "code")) 

#this will join region_defs to catch_clean 
#by = c("Region" = "code") this means that the left (Catch_clean file column name) hand column is equal to the column of the name "code"

head(catch_joined)


```
you can also list the pairs to join by multiple columns. 


```{r}
dates_df <- data.frame(date = c("5/24/1930",
                                "5/25/1930",
                                "5/26/1930",
                                "5/27/1930",
                                "5/28/1930"),
                       stringsAsFactors = FALSE)

```
 

###separating the components of one column into other columns using the separate function. Example: you can break down a "Date" column into Year, month and day columns 

```{r}

dates_split <- dates_df %>% 
  separate(date, into = c("month", "day", "year"), sep = "/", remove = F)

#into is the what names of columns you want to split into 
#Sep will be by what character you want to separate by 
#remove = F will leave the original column in place 

head(dates_split)

```

```{r}
dates_unite <- dates_split %>% 
  unite(col = date_iso, year,month,day, sep = "-") #columns that you want to unite, want these to be joined by separated by a dash 

head(dates_unite)
```
### Formating timestamp - the best format for a date is YYYY/MM/DD 




# Misc. Functions 


